Color- and piece-blind chess

### 2019-03-09 chess/papers/blind.tex
Some conclusions and things

Fortunately, we know the
 - there exists a perfect matching (players playing legally)
 - we can tell if we found one

so you have a set of black and white games, same number of each.
goal is to get an even score, while using O(n) space.
So we have an invariant that each of the games-as-white is paired
with at least one compatible game-as-black (and vice versa). Of
course at the beginning all pairs are compatible, and then white
makes a move; the games are still compatible because any of these
moves can be played on the black side. We just need to make sure
that we don't play e.g. e4 in all the games, because we need enough
black opponents to be playing against d4s and Nf3s and so on.
So, after all the white players have made their moves, we have
those boards, and need to make moves for black. For each, just pick
a compatible board (i.e., something that represents a legal move)
and play it. When we do so, we update the board to a state like
``I made this move as white.'' (So at the very beginning, all boards
are in the starting position with like ``I made this move as black,''
of course not really, more like ``white opponent to move'').
This is easy on the first move because all choices (that are in
the state ``copied this move from white'').

Now we wait for black players to make their moves. (Note waiting
can be done in constant space. You just compare each board that
has ``black opponent to move'' with the actual board state; if
the move is compatible then you update it. But might you confuse
the identities of two games? It would need to be the case that
there exists board states $B_1$, $B_2$, $B$ where $B$ can
legally follow $B_1$ and $B_2$ but $B_1$ and $B_2$ are different.
Of course this is possible! $B$ could be like

and then we have $B_1$ and $B_2$ as

\chessboard[tinyboard, setfen=1k6/8/8/8/8/8/8/K7 b - -, showmover]
\chessboard[tinyboard, setfen=Rk6/8/8/8/8/8/8/K7 b - -, showmover]

And so if we misassign $B$ to $B_1$, and then later see $B'$:

We are stuck because $B'$ cannot follow $B_2$.

So if we're actually going to do it with this method, we would
need to be able to compute the bipartite matching looking at all
the constraints. There certainly exists one, and any one will do. 
(Actually what we want is called a ``perfect matching.'')
So the question is, can we compute a perfect matching in constant
space? I mean, it seems like not. How could we even {\em store}
the matching? Well, we don't actually need to store the matching
because it's not *literally* what we need; we just want to have
assigned the games according to a perfect match. So one approach
is like, pick a random assignment. If we get stuck, just start
over (so we would have to keep two board states, in order to
be able to ``undo'', but that is still constant) and commit
the whole thing if we succeed. We're actually quite likely to
succeed since ambiguities are pretty rare. So this may work
fine in practice. The main thing we achieve here is that we
don't need to store the matching itself between rounds; the
matching is ``expensive'' because it contains an index, which
is logarithmic size in the number of games. So, with this
method, we get $O(n)$ space, which is actually around $200 n$.

For each pair, we store the state of the board $B$
as last we saw it in $G_w$ or $G_b$.

both Walice and Bob have made their moves (that is, it is white's turn
in $G_w$ and

$G_i$, we look up which pair $p$ it is, and look at $B_p$. This board

Note that

Computers can still impress us and each other with computational chess
feats other than playing. For example, computers are largely concerned
with filling up their memories with efficiently encoded data.

compactly storing a position. constant state

I think a version I liked was 64 bits (this gives the positions of all
pieces),\footnote{(We can also do a little better. Note that only up
  to 32 of the 64 bits can be set, so you could represent
  $\binom{60}{32}$ in $\sim 61$ bits and use e.g. a third king to
  indicate that there are no more actually occupied squares if fewer
  than 32 pieces are present. However, this gets vastly more
  complicated for only 3 bits of savings.)} followed by 32 records
giving the piece values (there are 6: pawn, knight, bishop, rook,
queen, king) and colors. To pick between the 12 possibilities we need
4 bits, and there is some slack here of course. anyway, that's 64 +
32*4 = 192 bits, plus 15 bits for other state. Note we could use the
c\_rook trick to save 4 bits for castling rights (14 possibilities),
not to mention a en\_passantable pawn (15) and a king whose turn it is
(even 16). Clear slack here, but that would leave only the 50-move
rule at 7 bits, for a total of 199. Wikipedia says the huffman
approach is ``maximum of 204 bits, and often much less.'' but the
article also contains many bugs, like the misconception that there can
only be four rooks.

So now comes the main idea of the paper, which is also spoilered in the
very clear paper title. What if we represented boards {\em only} as the
64-bit mask telling us what squares are occupied? The encoding is very
lossy, of course, but it often contains enough information to deduce
the state of the board, or ...

% 
% % make -j eval-unblinder.exe && ./eval-unblinder.exe net.val
% .!Evaluating model net.val...
% Loaded 50000 positions in 122.59s
% Loaded model in 0.10s
% Ran eval in 2.04s
% Reading [net.val]
% net.val: 4 layers.
% net.val: num nodes: 64 1024 12288 567 837
% net.val: indices per node/fns: 49 LEAKY_RELU 36 LEAKY_RELU 127 LEAKY_RELU 78 LEAKY_RELU
% Read from net.val.
% Invert index:
% Check it:
% ModelInfo [339885 rounds, 223224128 examples]
% Over 50000 positions:
%   9584 exactly correct (19.17%)
%   161166 piece mistakes (3.22/pos)
%   1630 castling mistakes (0.03/pos)
%   19014 move mistakes (0.38/pos)
%   181810 total mistakes (3.64/pos)
% 
% % $ make -j eval-unblinder.exe && ./eval-unblinder.exe net-before-vacuum.val
% make: 'eval-unblinder.exe' is up to date.
% Evaluating model net-before-vacuum.val...
% Loaded 50000 positions in 118.80s
% Loaded model in 3.99s
% Ran eval in 332.10s
% Reading [net-before-vacuum.val]
% net-before-vacuum.val: 4 layers.
% net-before-vacuum.val: num nodes: 64 1024 12288 567 837
% net-before-vacuum.val: indices per node/fns: 64 LEAKY_RELU 1024 LEAKY_RELU 12288 LEAKY_RELU 567 LEAKY_RELU
% Read from net-before-vacuum.val.
% Invert index:
% Check it:
% ModelInfo [146999 rounds, 9407936 examples]
% Over 50000 positions:
%   10601 exactly correct (21.20%)
%   156052 piece mistakes (3.12/pos)
%   1542 castling mistakes (0.03/pos)
%   18818 move mistakes (0.38/pos)
%   176412 total mistakes (3.53/pos)
% 
% 
% 
% without concurrent processes:
% net.val: 1.61s  (something like 1035/sec/core)
% net-before-vacuum: 229.85s (something like 7/sec/core)
% 
% 
% single_kings net.val:
% Over 50000 positions:
%   9642 exactly correct (19.28%)
%   162804 piece mistakes (3.26/pos)
%   1633 castling mistakes (0.03/pos)
%   19014 move mistakes (0.38/pos)
%   183451 total mistakes (3.67/pos)
% 
% It's actually kind of heartening that it performs worse with this approach,
% though we are more likely to get the whole board correct.
% 
% (cite #kingme)
%

% all bits set to 1 (with single_kings)
% (without single_kings, only difference is that the bottom right becomes a
% rook; no white queen)
\includegraphics[width=0.9 \linewidth]{blind-allon}

>>>>

Fortunately, we know that there exists a perfect matching (assuming
all players are playing legal moves) and we can tell if we found
one (i.e., we didn't get stuck). So, one strategy that works is
to choose randomly when there is some ambiguity, and start over
from the beginning if we ever get stuck. In practice this will
be pretty efficient, since convergent board states are unusual.
We only need a single pseudorandom pool for the entire process,
so it can be O(n) bits; this seems easily enough to generate all
possible permutations of $n/2$ items. Even $n^{2n}$ grows much
faster than $n!$. If we don't like the random approach, I believe
it is also possible to compute \verb+next_permutation+ in
constant space; so we can just explicitly try all orderings for $B'$
(this takes exponential time).

Once we have paired up each $B$ and $B'$, we simply compute the
move (which we now know exists) and play it in that game. We
then wait for the black opponents to play their moves, copy the
resulting board states into our vector and repeat the above
process (but flipping ``black'' and ``white'').

Although this is more involved than the previous approach, and may
take exponential time, it allows us to play against $n$ simultaneous
opponents using $O(n)$ space!\footnote{ 
%
If the games in the exhibition were presented to us in some way that
allowed us to visit them in the same order multiple times, we could
perhaps get away with {\em only} the random pool? XXX }

\subsection{Board representations} \label{sec:representation}

The actual space used per game is primarily the representation of
a chess position, plus a few bits for bookkeeping. So, representing
boards compactly gives us a way to increase the number of simultaneous
games we can play for a given storage size.

Mainly, we need to store the pieces and their locations. There are a
few other bits, like whose turn it is (1 bit), whether each player can
still castle king- and queen-side (4 bits), and whether and where an
en passant capture is possible (4 bits).\footnote{
%
Technically, we need to store a lot of additional information with the
board in order to completely implement the rules of
chess.\cite{fiderules} The trickiest of these involve the rules for
draw by repetition, which make reference to the history of the game
(See Footnote~1 in {\it Survival in chessland}\cite{survival} and seem
to require storing all previous board states. Fortunately, if we are
being this picky, then we also know that the length of a chess game is
bounded by a constant: Rule 9.6.2 ends the game in a draw if both
players have made 75 moves without a pawn move or
capture,\footnotemark\ so it suffices to store the $75\times 2$ most recent
(half-)moves. This sucks so most people don't do it (for example, FEN
notation only gives the number of such moves, and so cannot implement
the draw by repetition). On the other hand, if we insist, then this
may give us a simpler route to a constant-space exhibition, since the
$B \stepsto{m} B'$ relation is probably reversible with such
information.}\footnotetext{These are two types of moves that make it impossible
  to formally repeat a position that preceded them. Castling also has
  this property, but doesn't count because it is a ``secret move.''}

With 64 squares, six pieces for each of the two sides, plus the empty
square, a straightforward representation of the board uses $64 \times 4=256$
bits. The Thursd'z Institute considered more compact
representations\cite{representations}; one nice choice works as follows:

Start with a single 64-bit mask which indicates, for each square on
the board, whether it contains any piece. Note that there can only be
up to 32 pieces on the board. To tell what these pieces are, we then
follow with 32 four-bit records; these indicate the piece's color and
type.\footnote{Since only 32 of the 64 bits can be set, you could do
  slightly better by representing $\binom{60}{32}$ in $\sim 61$ bits.
  When fewer than 32 squares are occupied, we can use a record
  containing e.g. a third king (otherwise impossible) to indicate that
  we should ignore the remaining bits. However, this gets vastly more
  complicated for only 3 bits of savings.} 
%
With the 9 bits of extra state above, this totals $64+32 \times 4+9=201$
bits. There is some slack in the representation, because there are
only 12 actual possibilities for a piece but we can indicate 16 with 4
bits. It is great to abuse this slack to save bits; for example, we
can store a new type of rook that is still able to castle (it can only
be in the corners and thus also indicates its own color), eliminating
the 4 castling bits. We can similarly introduce an en-passantable
pawn, saving the 4 bits of en passant state; this piece can only be in
ranks 4 or 5, so it also indicates its color. We can also have a
``king whose turn it is'' for each side, saving the side-to-move bit.
This totals a nice round $64+32 \times 4=192$ bits.\footnote{ Since this is
  SIGBOVIK, I am freed from the burden of comparing related work. I
  did however read the rather bad Wikipedia article on the
  topic\cite{wikipediaboard} which describes a Huffman-based encoding
  that uses a ``maximum of 204 bits, and often much less.'' This also
  includes a 7-bit 50-move counter (but you really need to implement a
  75-move counter; 50 moves only allow a player to {\em claim} a draw)
  so should be compared as $197$ bits. But the article also contains
  many bugs, like the misconception that there can only be four total
  rooks (pawns are allowed to promote to rook). So the approach
  described here is both {\em more efficient} and {\em more correct}.}
%
This would allow approximately an 11 billion-game simultaneous
exhibition in RAM on my desktop computer.

\medskip
So now comes the main idea of the paper, which is also spoilered in the
very clear paper title. What if we represented boards {\em only} as the
64-bit mask telling us what squares are occupied? The encoding is very
lossy, of course, but it often contains enough information to deduce
the state of the board. For example, can you tell what position this is?

\chessboard[setfen=8/8/8/8/8/8/8/8,
  showmover=false, 
  smallboard,  
  color=gray,
  strokeopacity=0,
  markstyle={[fill]circle},
  markfields={a1,b1,c1,d1,e1,f1,g1,h1,a2,b2,c2,d2,e4,f2,g2,h2,a7,b7,c7,d7,e7,f7,g7,h7,a8,b8,c8,d8,e8,f8,g8,h8
}]

Correct! It is

\begin{tabular}{p{1.85in}p{1.4in}}
\chessboard[setfen=nrbqkbrn/pppppppp/8/8/4P3/8/PPPP1PPP/NRBBNRQK b - - 34 19, showmover=false, smallboard] & {
\vspace{-1.65in} \small after 1. Nf3 Nf6 2. e4 Ng4 3. Bc4 Ne5 4. O-O Ng6
5. Kh1 Rg8 6. Nc3 Nh8 7. Qe2 Nc6 8. Rb1 Rb8
9. Nb5 Nb4 10. Rd1 Nd5 11. Qf1 Nb6 12. Qg1 Na8
13. Nbd4 Nb6 14. Rf1 Na8 15. Be2 Ng6 16. Bd1 Nh8
17. Nb3 Ng6 18. Ne1 Nh8 19. Na1}
\end{tabular} \vspace{1em}

\section{Color- and piece-blind chess}

So this is kind of like blindfold chess, but for computers! Instead of
being blind to the board, and only relying on our memory (trivial for
computers), we'll only be able to see where pieces are positioned on
the board, but not what type or color they are. Of course, we also
have to prohibit the computer from simply remembering the board, so
the algorithm must be stateless. Specifically, we want a function
\[
  {\tt makemove} \;:\; {\tt uint64} \rightarrow {\tt move\ list}
\]
that makes a move from a single board position, represented just as 64
bits. This is a single move; the move list represents our preference
for the move to make in descending order, and we commit to the first
move that is actually legal. It does not work well to insist that the
function return a single move, as it will often be the case that the
board is misinterpreted and a move is in fact illegal; forcing
forfeit\footnote{FIDE rules state that the {\em second} attempt at an
  illegal move results in forfeit (7.5.5).} would mean that almost all
games end in forfeit, which is boring. On the other hand, allowing the
function to try again upon learning a move is illegal would allow it
to interactively ``interrogate'' the board state
somewhat.\footnote{Similar to Kriegspiel~\cite{wikipediakriegspiel}, although
  in that game at least one's own pieces are known!} This seems
counter to the spirit of color- and piece-blind chess, so we instead
require the function to rank all moves ahead of time.

\section{Unblinding the board}

I went about this by building a function that ``unblinds'' a board; it
has type
\[
{\tt unblind} \;:\; {\tt uint64} \rightarrow {\tt position}
\]

This function is natural for machine learning. It is easy to generate
copious training data from actual games by simply blinding positions
into their corresponding 64-bit numbers; I just randomly sampled 100
million positions from Feburary~2017 on lichess.org.

I repurposed the custom neural network code from my seminal paper {\it
  Red i removal with artificial retinal networks}~\cite{redi} after
discovering that artificial retinal networks are actually {\em
  isomorphic} to neural networks. The main advantage of this code is
that it allows for sparse networks, but the real reason to use it is
that I would rather spend dozens of hours debugging my own code, pay a
larger electric bill, and get worse results in the end, than to spend
a short while trying to get someone else's probably-very-good neural
network training package to compile.

The structure of the network is as follows. The input layer is 64
nodes, one for each of the 64 bits, with each node set to either {\tt
  1.0f} or {\tt 0.0f}. Three hidden layers of size 1024, 12288, and 567
do the neural magic. The output layer is 837 nodes; the bulk of which
is a ``one-hot'' representation of the predictions for the 64 squares,
each with 13 possible contents (black or white, six pieces, or empty).
This is $64 \times 13 = 832$ nodes. Then four nodes to predict the
four castling bits, and one to predict the current side to move. This
model does not predict the possibility for {\em en passant} capture,
nor move counters or position repetition. This will not be its main
source of disadvantage!

I trained the network in two phases, first starting with a densely
connected one (model size 160 MB), and then after I get fed up with
how slow training was, a ``vacuumed'' version of the network where I
removed edges with low absolute weights (model size 5 MB) to continue
training. Removing edges based on an absolute weight threshold is
very unprincipled (since a downstream edge can magnify a contribution
arbitrarily) but I did it anyway.

Everything was trained on a single GPU, though a fairly decent one in
2018, the EVGA GeForce GTX 1080 FTW (really), using OpenCL. Biases
were initialized to $0$ and weights with a Gaussian of mean $0$ and
standard deviation $0.025$. In the first phase, there were 64 examples
per round, and after vacuuming, 2048. The round learning rate
$\alpha_r$ started at $0.1$ and descended linearly to $0.002$ over
500,000 rounds; and the learning rate when updating weights (for each
example) $\alpha$ is $\alpha_r / \verb+examples_per_round+$. In no way
am I recommending these parameters. Fiddling with the parameters to
make it do its black magic or alternately carom off to a sea of NaNs
or zeroes is for sure the worst part of neural networks. In fact, I
initially started with the classical sigmoid transfer function, but
``upgraded'' to the ``leaky rectified linear'' function
%
$$(p < 0) \,?\, p \times 0.01 : p$$
%
after getting fed up with sigmoid weights caroming off (see
``vanishing gradient problem'' and/or ``exploding gradient problem'').
The final model was trained over 339,885 rounds on 223 million
examples. It did not appear to show signs of improving for
several days before I terminated it.

\subsection{Statistical evaluation}

The unblinding component can be evaluated on its own, by running it on
examples sampled independently of the training set. The model outputs
a score for each possible contents of each square; we simply
discretize to the highest-scoring one (same too for the predicted
castling state and turn). Over 50,000 examples, these were the
results:

9,584 predicted boards were exactly correct (19.17\%). There were a
total of 161,166 piece mistakes, which is an average of 3.22 per
board. This is wherever we predict a square's contents incorrectly.
There were only 1630 castling status mistakes, an average of 0.03 per
board (there can be up to four mistakes per board). This is probably
because when the king and rook are in their starting squares, castling
is almost always still allowed. In 19,014 boards, we mispredicted
whose move it is (38\%). This appears to be the most difficult thing
to predict, which is not surprising.\footnote{ Prior to ``vacuuming'',
  the 160 MB network actually performed slightly better than the final
  5 MB network, with 21.20\% of boards exactly correct, and an average
  of 3.12 mistakes per board. This suggests that the model may only be
  doing a limited amount of generalization, instead mostly memorizing
  board positions. Representing the 223 million examples seen exactly
  (using our best board representation described in
  Section~\ref{sec:representation}) would take 42.8 GB, so at 5 MB at
  least the data is represented compactly, if also rather lossily.}

\subsection{Subjective evaluation}

The unblinder {\em must} make mistakes since the 64-bit representation
is ambiguous. Subjectively, the unblinder makes reasonable mistakes.
It is excellent at common openings, usually getting these exactly
correct. On the other hand, it is fairly bad at sparse endgames, where
it is difficult to tell a pawn from a rook from a king. It is terrible
at unlikely positions that can be confused for likely ones. If you are
playing against it and know how it works, it is easy to trick it by
doing something like capturing one of its starting-position pawns with
your queen; nobody does this in real games (because the queen can be
immediately recaptured), so the square is predicted as a pawn and the
queen ``disappears'' to the unblinder (Figure~\ref{fig:trick}). Having
an invisible queen in your camp is of course very dangerous. Resolving
ambiguities in favor of more likely positions is the right thing for
the model to do, so this is just an inherent flaw with the
decomposition of the problem. There are some ways we can account for
this (Section~\ref{sec:spycheck}).

\begin{figure}[ht]
XXX actual example
\caption{The white queen ``disappears'' after QxWhatever, because
  unblinding predicts it to be one of black's own pawns, which is far
  more likely in that square.} \label{fig:trick}
\end{figure}

A few things are distinctly disappointing about its performance. Even
outside of ``likely'' positions, it usually predicts that pieces on
black's side of the board are black, and vice versa
(Figure~\ref{fig:allon}). This makes sense, but suggests serious
limitation on using the prediction to play chess. Less forgivably, it
sometimes predicts more than one king per side (or zero), which is
always wrong. Actually, an early version had this problem in spades,
usually frequently predicting two or three kings. Upon debugging, I
had simply made the noob mistake of printing both King and Knight with
the letter ``K.'' Ha! It often predicts the ``wrong'' number of
bishops (etc.), or places them on the same color. This is technically
possible through promotion, but usually a bad guess, since promotion
is relatively rare, and moreso promotion to a piece other than a
queen. An approach that might handle this better (but may have
different downsides) would be instead to predict the ``fates'' of the
32 initial pieces~\cite{survival}. The fate of a pawn includes what if
any piece it has promoted to, but no other pieces can. This would {\em
  require} that the model only predict a single location for each
king, among other things. However, this would require a much larger
output layer (32 pieces can move to 64 squares, plus promotion) and it
is not always clear how to interpret its value into a position (for
example, if two pieces are predicted strongly to be on the same
square).

\begin{figure}[ht]
\includegraphics[width=0.9 \linewidth]{blind-allon}
\caption{What the model predicts (in single-king mode;
  Section~\ref{sec:using}) for {\tt 0xFFFFFFFFFFFFFFFF}, the board
  with all bits set. This is an impossible position, but it gives some
  idea of the model's biases for each square. Notably, most of the pieces
  on each half of the board have a single color. This makes sense, but
  also suggests substantial limitation. When single-king mode is off,
  the bottom right king is predicted as a white rook.} \label{fig:allon}
\end{figure}

\section{Playing blind}

Once we have predicated a board state, we can play with it. The
simplest way to do this is to use a strong chess engine to pick a move
for the position. Here, I use Stockfish with a node budget of 1
million nodes, which takes about 1 second of CPU time per move on my
computer; henceforth {\sf stockfish1m}.

There are some complications:
\begin{itemize}
\item Frequently, the unblinded board will not match the true
  position, and Stockfish will choose a move that is illegal. So, as discussed
  before, we actually return a prioritzed list of moves. For this first
  experiment, we just return the move Stockfish recommends, followed by all
  other moves ordered randomly.
\item Stockfish is a very strong engine, and in my opinion the code is generally
  good, but it is very sensitive to bad FEN (the notation used to give it a
  position to evaluate) strings. Given a bad string, like one that says castling
  is possible when the king isn't in its home square, often crashes the engine.
  So we need to make sure to actually pass valid positions. I accomplish this
  by making the following modifications to the predicted board:
  \begin{itemize}
  \item If a castling bit is set, but castling is not possible due to the position
    of the king or rook, I clear the bit.
  \item Set the side-to-move to be the correct actual value. This uses
    the unblinded state, so is superficially cheating. But note that
    if we get the side wrong, then Stockfish's move will always be
    illegal: Moves are specified as a source and destination
    square,\footnote{Plus promotion piece. Castling is represented as
      a two-square move of the king.} and so the source square of
    Stockfish's move would always be a piece of the wrong piece's
    color. So this is equivalent to (but twice as efficient as)
    running stockfish twice, one for each side, and prioritizing the
    predicted side's move first.
  \end{itemize}
  This won't fix all positions, for example, if the white and black king are
  adjacent to one another in mutual check. If an irreparable problem is detected,
  then I just return a uniformly random move list.
\end{itemize}

It is easy to beat this chess engine, by tricking it as in
Figure~\ref{fig:trick}, although this involves unnatural moves, so it
may only apply if you know how it works.
% XXX example game?
Measuring how well it plays in an absolute sense is a subject of some
interest, so I wrote a separate paper about that~\cite{eloworld}. This
algorithm, called \verb+blind_yolo+, had an Elo World score of XXX.

\subsection{We three kings}

When evaluating the first version I found that it was predicting a
disappointingly high number of illegal positions in practice, which
was causing us to fall back on making random moves, which is mostly
boring. The second version reduces the rate of illegal positions due
to too many or too few kings~\cite{kingme}.

The model predicts a score for each type of piece in each square, and
we do not have to necessarily interpret it by always taking the
highest-scoring piece. This version first finds the two squares with
the highest scores for the white king, and same for the black king. We
take two in case the same square is predicted for both. Then this
square gets one of the kings (whichever has higher score) and the
other king goes in the highest-scoring unclaimed square. The rest of
the squares get the highest-scoring prediction as before, but we never
predict kings for them.

This change just affects the unblinding process, so we can directly
evaluate its accuracy. It gets 19.28\% of positions exactly correct
(slightly better), with an average of 3.26 piece mistakes per position
(slightly worse). This is expected; we exchange local mistakes (each
was trained independently to minimize its local error) for global
correctness (which is not taken into account at all during training).

This version, called \verb+blind_kings+, performs consistently better
than \verb+blind_yolo+ (63 wins, 45 losses, 412 draws).
% XXX could run a lot more games between these two.
% as white: 31/24/207, as black 21/32/205
It had an Elo World score of XXX.

\subsection{Spy check} \label{sec:spycheck}

Say \verb+blind_kings+ is playing as black; it remains easy to fool it
by moving black pieces into white's camp, since they are usually then
predicted to be white pieces. We can defend against this somewhat.
Since it is illegal to capture one's own piece, there is little risk
in trying; if it is indeed our own piece then the move will be
rejected, and if it is not our piece, then capturing is good for two
reasons: One, we capture a piece, and two, we avoid having
Stockfish make a move in this incorrectly predicted board. (Of course
there are many reasons why eagerly capturing a piece can be a bad
idea, but at this level of play, an edge in material is likely worth
it.)

There is one subtlety here. Above we argued that it was safe to use
the actual side-to-move instead of the predicted one; but here it
would not be equivalent to do so. Instead, we first prioritize all
apparent spy-check moves where the predicted source piece matches the
predicted side-to-move, then we try the opposite. (Ties are broken by
preferring to capture with a lower-value predicted piece, and then
randomly.) Due to this, there is some additional chance that we end up
making an especially dumb move because we both mispredicted the
side-to-move and the identity of some pieces.

(XXX evaluation here)

\subsection{Future work}

The predicted board often expresses uncertainty about some squares,
which could be thought of as probabilities. A principled improvement
would be to try to find moves that are good in expectation, that is,
integrated all possible boards in proportion to their predicted
probability. A good approximation might be had by sampling a bunch
of boards according to the predicted distribution, and then using
Stockfish to score the top $k$ moves for each; we can then order
moves by their expected score. Unfortunately, it is not easy to
efficiently get Stockfish to generate scored moves for $k \neq 1$.
Even with $k = 1$, this approach would be slow, taking about a
second for each (distinct) sampled board. So I did not try it,
at least not before submitting this paper.

\medskip
I initially considered trying to solve this whole problem with neural
networks. The current best known engine in the world (AlphaZero) at
least {\em uses} a neural network.
% XXX is the whole thing just a network? I guessed it's probably
% alpha-beta plus a NN evaluation function?
The biggest advantage would be that it would naturally be able to
consider multiple moves under uncertainty about the board state, as
just discussed, without any particular extra logic. My plan was to
make a multiple different components that coud be evaluated
separately, starting with the unblinder described, followed by a unit
that predicts legal moves, and then a unit that takes these two (and
also the 64-bit blinded mask if it likes) and scores each move.
Training a reasonable unblinder took longer than I expected, and the
legal move predictor never really worked that well,
% (XXX some evaluation of it)
so I just settled for basing it off the single unblinder unit. Can you
do better?

\section{Conclusion}
% 
% - How do the solutions compare to other chess algorithms that
%   could be applied in the blind? Random works, also lexicographic,
%   any others? Most are inspecting the board state, obviously.
% - Could estimate the upper bound for accuracy by blinding every
%   game in the database, and doing a ``lookup''-based version of
%   the eval.

I would like to thank the little people (pawns) and the anonymous
referrees.

% % make -j eval-unblinder.exe && ./eval-unblinder.exe net.val
% .!Evaluating model net.val...
% Loaded 50000 positions in 122.59s
% Loaded model in 0.10s
% Ran eval in 2.04s
% Reading [net.val]
% net.val: 4 layers.
% net.val: num nodes: 64 1024 12288 567 837
% net.val: indices per node/fns: 49 LEAKY_RELU 36 LEAKY_RELU 127 LEAKY_RELU 78 LEAKY_RELU
% Read from net.val.
% Invert index:
% Check it:
% ModelInfo [339885 rounds, 223224128 examples]
% Over 50000 positions:
%   9584 exactly correct (19.17%)
%   161166 piece mistakes (3.22/pos)
%   1630 castling mistakes (0.03/pos)
%   19014 move mistakes (0.38/pos)
%   181810 total mistakes (3.64/pos)
% 
% % $ make -j eval-unblinder.exe && ./eval-unblinder.exe net-before-vacuum.val
% make: 'eval-unblinder.exe' is up to date.
% Evaluating model net-before-vacuum.val...
% Loaded 50000 positions in 118.80s
% Loaded model in 3.99s
% Ran eval in 332.10s
% Reading [net-before-vacuum.val]
% net-before-vacuum.val: 4 layers.
% net-before-vacuum.val: num nodes: 64 1024 12288 567 837
% net-before-vacuum.val: indices per node/fns: 64 LEAKY_RELU 1024 LEAKY_RELU 12288 LEAKY_RELU 567 LEAKY_RELU
% Read from net-before-vacuum.val.
% Invert index:
% Check it:
% ModelInfo [146999 rounds, 9407936 examples]
% Over 50000 positions:
%   10601 exactly correct (21.20%)
%   156052 piece mistakes (3.12/pos)
%   1542 castling mistakes (0.03/pos)
%   18818 move mistakes (0.38/pos)
%   176412 total mistakes (3.53/pos)
% 
% 
% 
% without concurrent processes:
% net.val: 1.61s  (something like 1035/sec/core)
% net-before-vacuum: 229.85s (something like 7/sec/core)
% 
% 
% single_kings net.val:
% Over 50000 positions:
%   9642 exactly correct (19.28%)
%   162804 piece mistakes (3.26/pos)
%   1633 castling mistakes (0.03/pos)
%   19014 move mistakes (0.38/pos)
%   183451 total mistakes (3.67/pos)
